---
title: "PA1_template"
output: html_document
---

Reading the file

```{r}
steps <- read.csv("activity.csv")
stepsmod <- steps[complete.cases(steps),]
```

Calculating the total number of steps per day

```{r}
library(dplyr)
stepstotal <- stepsmod %>% group_by(date) %>% summarise_each(funs(sum))
stepstotal
```

Histogram of total number of steps per day

```{r}
hist(stepstotal$steps, main = "Histogram of total number of steps", xlab="No of steps", ylab = "Frequency")
```

Mean and median for the total number of steps per day

```{r}
stepsmean <- stepsmod %>% group_by(date) %>% summarise_each(funs(mean))
stepsmean

stepsmedian <- stepsmod %>% group_by(date) %>% summarise_each(funs(median))
stepsmedian
```
5-minute inetrval plot

```{r}
stepsmean2 <- stepsmod %>% group_by(interval) %>% summarise_each(funs(mean))
plot(stepsmean2$interval,stepsmean2$steps,type = "l", xlab = "5-minute interval", ylab = "Average number of steps taken")
```


Calculating the number of NAs

```{r computena}
na <- sum(is.na(steps$steps))
```

Total number of rows with NAs is `r na`

Replacing NAs with mean

```{r}
steps[is.na(steps)] <- mean(stepsmod$steps)
```

Creating histogram and calculating mean and median for the new dataset

```{r}
stepstotal2 <- steps %>% group_by(date) %>% summarise_each(funs(sum))

hist(stepstotal2$steps, main = "Histogram of total number of steps", xlab="No of steps", ylab = "Frequency")

stepsmean2 <- steps %>% group_by(date) %>% summarise_each(funs(mean))
stepsmean2

stepsmedian2 <- steps %>% group_by(date) %>% summarise_each(funs(median))
stepsmedian2
```

Adding weekdays and weekends

```{r}
library(lubridate)
stepsdays <- mutate(steps,daytype = ifelse(wday(steps$date) == (1 | 7),"Weekend","weekday"))
stepsdays$daytype <- factor(stepsdays$daytype)
```

Plotting by weekdays and weekends

```{r}
stepsdaysmean <- stepsdays %>% group_by(interval,daytype) %>% summarise_each(funs(mean))

par(mfrow = c(2,1), mar = c(4, 4, 2, 1))
with(subset(stepsdaysmean, daytype == "weekday"), plot(stepsdaysmean$interval,stepsdaysmean$steps, type = "l", main = "Weekday", xlab = "5-minute Interval", ylab = "Average number of steps"))
with(subset(stepsdaysmean, daytype == "weekend"), plot(stepsdaysmean$interval,stepsdaysmean$steps, type = "l", main = "Weekend", xlab = "5-minute Interval", ylab = "Average number of steps"))
```